### 2015-01-07 doc/thesis/evaluation.tex
Minor improvements to the text about other projects

In this section we compare Typed Lua with other projects and type systems.
We split this section into two parts: the first compares Typed Lua
with other Lua projects, while the second compares Typed Lua with
other type systems that are not related to Lua.

@@

The table type from Tidal Lock can only represent records, that is,
it cannot describe hash tables and arrays yet, though we can refine them.
Tidal Lock also includes field types to describe the type of the fields
of a table type.
The field types describe if a table field is mutable or imutable
in a table type.
Field types are the feature that allow the refinement of table types in
Tidal Lock.

@@

\subsection{Other type systems}

@@

The type system of Typed Racket introduced \emph{occurrence typing}
\citep{tobin-hochstadt2010ltu}.
It is a novel way to use type predicates in control flow statements
to refine union types.
The type system is rich and includes basic types, function types,
union types, recursive types, and structure types.
Function types can return multiple values.
Structure types are similar to our interfaces, as they describe
record types.
The type system is also structural and based on subtyping.
It also includes the dynamic type \texttt{Any}, which is the top
type in the system.
Typed Racket also supports polymorphic functions and data structures.

@@

The type system of Typed Clojure is rich.
It includes polymorphic function types, union types, intersection types,
lists, vectors, maps, sets, and recursive types.
Function types can also have rest parameters, which are similar
to our variadic types, but can only appear on the input parameter
of function types.
In fact, its function types cannot return multiple results.
It also uses ocurrence typing to allow control flow statements to
refine union types.
The type system is also structural and based on subtyping.

Dart \citep{dart} is a new class-based object-oriented programming
language.
It includes optional type annotations and compiles to JavaScript.
The type system of Dart is nomial and includes base types,
function types, lists, and maps.
It also supports generics, and the programmer can define
generic functions, lists, and maps.

@@

TypeScript \citep{typescript} is a JavaScript extension
that includes optional type annotations and class-based
object-oriented programming.
It also uses local type inference to deduce the type
of unannotated expressions.
The type system of TypeScript is structural and based
on subtyping.
It includes the dynamic type, primitive types, function
types, array types, tuple types, and recursive types.
TypeScript also support generics.

@@

@@

Gradualtalk inserts run-time checks that ensure dynamically
typed code does not violate statically typed code.
\citet{allende2013cis} perform a carefull evaluation about
cast insertion on Gradualtalk.
They report that usually cast insertions impact on execution
performance.

Reticulated Python \citep{vitousek2014deg} is a Python compiler
that supports gradual typing.
The type system is structural and based on subtyping.
It includes base types, the dynamic type, list types,
dictionary types, tuple types, function types, set types,
object types, class types, and recursive types.
It includes class and object types to differentiate the
type of class declarations and instances.
It also uses local type inference.
Besides static type checking, Reticulated Python also introduces
three different approaches for inserting run-time assertions.

Hack \citep{hack} is a new programming language that combines
static and dynamic type in PHP code.
Its type system includes generics, nullable types, collections,
and function types.

@@

>>>>

In this section we compare Typed Lua with other projects.
We split this section into two parts: the first compares Typed Lua
with other Lua projects, while the second compares Typed Lua with
other projects that are not related to Lua.

@@

The table type from Tidal Lock can only represent records, that is,
it cannot describe hash tables and arrays yet, though we can refine them.
Tidal Lock also includes field types to describe the type of the fields
of a table type.
The field types describe if a table field is mutable or immutable
in a table type.
Field types are the feature that allow the refinement of table types in
Tidal Lock.

@@

Luacheck \citep{luacheck} is a tool that performs static analysis on Lua code.
It can flag access to undeclared globals and unused local variables,
but it does not perform static type checking.

\subsection{Other projects}

@@

The main feature of Typed Racket's type system is \emph{occurrence typing}
\citep{tobin-hochstadt2010ltu}.
It is a novel way to use type predicates in control flow statements
to refine union types.
Occurrence typing is not sound in the presence of mutation.
As these kinds of checks are common in other languages, related systems
have appeared \citep{guha2011tlc,winther2011gtp,pearce2013ccf}.

The type system of Typed Racket also includes function types, recursive
types, and structure types.
Its function types also handle multiple return values, and there is
also a way to describe function types that have optional arguments.
Its structure types are similar to our interfaces, as they describe record types.
The type system is also structural and based on subtyping.
It also includes the dynamic type \texttt{Any}, which is the top type in the system.
Typed Racket also supports polymorphic functions and data structures.

@@

The type system of Typed Clojure includes polymorphic function types,
union types, intersection types, lists, vectors, maps, sets, and recursive types.
Function types can also have rest parameters, which are similar
to our variadic types, but can only appear on the input parameter
of function types.
In fact, its function types cannot return multiple results.
It also uses occurrence typing to allow control flow statements to
refine union types.
The type system is also structural and based on subtyping.

Dart \citep{dart} is a new class-based object-oriented programming
language.
It includes optional type annotations and compiles to JavaScript.
The type system of Dart is nominal and includes base types,
function types, lists, and maps.
It also supports generics, and the programmer can define
generic functions, lists, and maps.

@@

TypeScript \citep{typescript} is a JavaScript extension
that includes optional type annotations and class-based
object-oriented programming.
It also uses local type inference to deduce the type
of unannotated expressions.
The type system of TypeScript is structural, based
on subtyping, and supports generics.
It includes the dynamic type, primitive types, function
types, array types, tuple types, recursive types, and
object types.
Unlike Typed Lua, TypeScript does not include union types,
and it uses arrays to represent variadic functions and
multiple return values.

@@

TeJaS \citep{lerner2013tejas} is a framework for the construction of
different type systems for JavaScript.
The authors created a base type system for JavaScript with
extensible typing rules that allow the experimentation of
different static analysis.
They used TeJaS to create a type system that simulates the
type system of TypeScript.

@@

Gradualtalk inserts run-time checks that ensure dynamically
typed code does not violate statically typed code.
\citet{allende2013cis} perform a carefull evaluation about
cast insertion on Gradualtalk.
They report that usually cast insertions impact on execution
performance, so Gradualtalk also has an option that allows
programmers to turn them off, downgrading Gradualtalk
to an optional type system.

Reticulated Python \citep{vitousek2014deg} is a Python compiler
that supports gradual typing.
The type system is structural and based on subtyping.
It includes base types, the dynamic type, list types,
dictionary types, tuple types, function types, set types,
object types, class types, and recursive types.
It includes class and object types to differentiate the
type of class declarations and instances, respectively.
It also uses local type inference.
Besides static type checking, Reticulated Python also introduces
three different approaches for inserting run-time assertions.

Hack \citep{hack} is a new programming language that runs on the
Hip Hop Virtual Machine (HHVM).
The HHVM is a virtual machine that executes Hack and PHP programs.
We can view Hack as an extension to PHP that combines static and
dynamic typing.
The type system of Hack includes generics, nullable types, collections,
and function types.

@@

Grace \citep{black2013sg} is an object-oriented language
with optional typing.
Grace is not a dynamically typed language that has been
extended with an optional type system, but a language
that has been designed from scratch to have both
static and dynamic typing.
\citet{homer2013modules} explores some useful patterns
that derive from Grace's use of objects as modules
and its brand of optional structural typing, which
can also be expressed with Typed Lua's modules as tables.
